### 2020-01-16 ChapterImplementation/chapterImplementation.tex
Updated: Visualiser section split into subsections. More details in the interactivity subsection and list of terminology

The visualiser connects to multiple servers, objects are simulated on server and those objects from all servers are replicated on the visualiser. These are referred to as replica objects. The visualiser uses RakNet's ReplicaManager, which is responsible for: the creation of replicas objects on both the server and client; communicating the state of objects from the server to the client, in order for the replica objects to be replicated; and the deletion of replica objects. Replica objects are replicated on the visualiser by the server sending the state (position, rotation) to the client, the replica representation on the client is then updated with the newly received state from the server. The object state is optimised to reduce processing and network overhead, by only sending updates of the position and rotation to the client, when they change on the server.

@@

The visualiser supports server-created replica objects, which are limited to pre-defined basic objects (cuboid, sphere \& capsule). A different render material per server is used to identify the owning server of each object. These server objects have pre-defined representations on the client. The server sends an identifier for the type of object, which is then created on the client. Using pre-defined objects means the whole geometry of the object does not need to be transferred to the client. This is useful for when there are a large number of identical objects as it avoids having to re-send the same geometry for each object. However, this method requires the client to know of these pre-defined types.

The visualiser also supports arbitrary client-created objects. Entire scenes can be created in the client at edit-time, in the Unreal Engine Editor, and then re-created on servers when connections to the servers are made. This works for objects with any mesh, i.e. the server does not need to know about the geometry of an object when the server is built and deployed, the client communicates all data required to create the physical object on the server. All objects marked for replication on the servers are sent to all servers, if an object exists outside of a server's region, the creation is skipped, meaning the replica object is created only on the server containing the replica object. The central point of the object is used, avoiding conflicts between servers when an object exists that intersects a server region boundary.

In addition to replica objects, static object can also be created by the client. As these objects are static, there is no need for their states to be replicated as they do not change over the course of the simulation. Static object work simply by the client sending the geometry, position and rotation details of the static object to all servers and the servers then re-create the static object. These static objects then remain unchanged throughout the simulation, requiring no updates to the state to be communicated to the client. Static objects are created on all servers, this could be optimised so only static objects within a server's region are created, however, AP can lead to clusters of objects being simulated outside of their host server's region, which may interact with static objects that do not exist on their host server. In addition, the cost of simulating static objects tends to be very low, as they do not require any updates as they are not dynamic. In the future, region boundaries may change; a server may become responsible for a static object previously residing in another server's region, which would then have to be communicated when the region boundaries changed, by all servers simulating all static objects, this avoids the need for extra communication when server boundaries change. With these trade-offs in mind, we have elected to create all static objects on all servers, when static objects are used in a simulation.

The visualiser also allows for interactivity, the creation of server-created objects can be triggered from the client, as well as allowing the user to reset or change scenes that are being simulated on the server.

Currently when server created objects are migrated between servers, the client destroys the replica and a new replica is created. This is because the simulated object is destroyed on one server and re-created on another. A better way of dealing with this, would be a procedure for handing over server-ownership of replicas, removing the need for replicas to be destroyed and recreated. This would reduce processing overhead on the client, as there is no longer a need to destroy and recreate objects when server migrations take place.

@@

>>>>

\subsection{Terminology}
Before discussing specific details of the visualiser, we must first define terms used.
\begin{itemize}
	\item The visualiser: The software application that is capable of connecting to AP servers and display replicas of objects on the servers.
	\item The editor: This refers to the Unreal Engine Editor. The version of the editor used for the visualiser is UE4.19.2
	\item Edit-time: When scenes are being edited in the Unreal Engine Editor, before the visualiser has been packaged.
	\item Run-time: When the visualiser is running
	\item Player: The in-visualiser representation of the user. Has a position and rotation. The camera for rendering is attached to the player
	\item Replica object: A representation displayed in the visualiser of an object being simulated on a server, which is intended to replicate the physical state of the object on the server.
\end{itemize}

\subsection{Overview}
The visualiser connects to multiple servers, objects are simulated on server and those objects from all servers are replicated on the visualiser. These are referred to as replica objects. The visualiser uses RakNet's ReplicaManager, which is responsible for: the creation of replicas objects on both the server and client; communicating the state of objects from the server to the client, in order for the replica objects to be replicated; and the deletion of replica objects. Replica objects are replicated on the visualiser by the server sending the state (position, rotation) to the client, the replica representation on the client is then updated with the newly received state from the server. The object state is optimised to reduce processing and network overhead, by only sending updates of the position and rotation to the client, when they change on the server.

@@

\subsection{Server-created replica objects}
The visualiser supports server-created replica objects, which are limited to pre-defined basic objects (cuboid, sphere \& capsule). A different render material per server is used to identify the owning server of each object. These server objects have pre-defined representations on the client. The server sends an identifier for the type of object, which is then created on the client. Using pre-defined objects means the whole geometry of the object does not need to be transferred to the client. This is useful for when there are a large number of identical objects as it avoids having to re-send the same geometry for each object. However, this method requires the client to know of these pre-defined types.

\subsection{Client-created replica objects}
The visualiser also supports arbitrary client-created objects. Entire scenes can be created in the client at edit-time, in the Unreal Engine Editor, and then re-created on servers when connections to the servers are made. This works for objects with any mesh, i.e. the server does not need to know about the geometry of an object when the server is built and deployed, the client communicates all data required to create the physical object on the server. All objects marked for replication on the servers are sent to all servers, if an object exists outside of a server's region, the creation is skipped, meaning the replica object is created only on the server containing the replica object. The central point of the object is used, avoiding conflicts between servers when an object exists that intersects a server region boundary.

\subsection{Client-created static objects}
In addition to replica objects, static object can also be created by the client. As these objects are static, there is no need for their states to be replicated as they do not change over the course of the simulation. Static object work simply by the client sending the geometry, position and rotation details of the static object to all servers and the servers then re-create the static object. These static objects then remain unchanged throughout the simulation, requiring no updates to the state to be communicated to the client. Static objects are created on all servers, this could be optimised so only static objects within a server's region are created, however, AP can lead to clusters of objects being simulated outside of their host server's region, which may interact with static objects that do not exist on their host server. In addition, the cost of simulating static objects tends to be very low, as they do not require any updates as they are not dynamic. In the future, region boundaries may change; a server may become responsible for a static object previously residing in another server's region, which would then have to be communicated when the region boundaries changed, by all servers simulating all static objects, this avoids the need for extra communication when server boundaries change. With these trade-offs in mind, we have elected to create all static objects on all servers, when static objects are used in a simulation.

\subsection{Interactivity}
The visualiser also allows for interactivity, the creation of server-created objects can be triggered from the client. This feature uses RakNet's Remote Procedure Calls (RPCs), which enables clients to execute procedures on the server, including any required parameters. The client sends a RPC (triggered by a key press) with the parameters of player position and forward direction. The RPC message is received by all servers, only the server which is responsible for the player position creates a server-created object at that position and the object is given a velocity in the direction of the player's direction. This allows the player to create new objects at run-time, which are then able to interact with all objects in the simulation.

The user is also able to reset or change scenes that are being simulated on the server. In the visualiser, the scene can be reset on a key press. This sends an RPC to all servers, which either resets all simulations or resets the current test scenario. As all objects are deleted on the servers, any replicas on the visualiser are deleted. Also available to the player is a menu which can be opened that allows the user to select a scene to switch to. This allows the user to change the scene in the visualiser at run-time while maintaining connections with all servers.

@@

\subsection{Future work}
Currently when server created objects are migrated between servers, the client destroys the replica and a new replica is created. This is because the simulated object is destroyed on one server and re-created on another. A better way of dealing with this, would be a procedure for handing over server-ownership of replicas, removing the need for replicas to be destroyed and recreated. This would reduce processing overhead on the client, as there is no longer a need to destroy and recreate objects when server migrations take place.
